\documentclass[MTech]{iitmdiss}

\usepackage{times}
\usepackage{epsf}
\usepackage{threeparttable}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{txfonts,pxfonts,amsfonts}
\usepackage{epsfig}
\usepackage{caption}
\usepackage{subfig}
\usepackage[dvips]{graphicx}
% \usepackage[square,numbers,sort]{natbib}
\usepackage[square]{natbib}
\usepackage[hypertex]{hyperref} % hyperlinks for references.
%\usepackage{algorithmic}
%\usepackage{algorithm}
\usepackage{listings}


%\include{commands}

% Strut macros for skipping spaces above and below text in tables. 
\def\abovestrut#1{\rule[0in]{0in}{#1}\ignorespaces}
\def\belowstrut#1{\rule[-#1]{0in}{#1}\ignorespaces}

\def\abovespace{\abovestrut{0.20in }}
\def\aroundspace{\abovestrut{0.20in}\belowstrut{0.10in}}
\def\belowspace{\belowstrut{0.10in}}
%%%%%%%%%%%%%%%%%%%%%%%%%


\def\thesistitle{Framework For Realizing Autoparallelism Through Automatic OpenMP Code Generation}

\def\thesisauthor{Raghesh A}


\begin{document}
\bibliographystyle{iitm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
% Title page

\title{\thesistitle}

\author{\thesisauthor}

\date{April 2011}
\department{Computer Science and Engineering}

%\nocite{*}
\begin{singlespace}
\maketitle 
\end{singlespace} 



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Certificate
\certificate

\vspace*{0.5in}

\noindent This is to certify that the thesis entitled {\bf {\thesistitle}}, 
submitted by {\bf {\thesisauthor}}, to the Indian Institute of Technology, 
Madras, for the award of the degree of {\bf Master of Technology}, 
is a bona fide record of the research work carried out by him under my
supervision. The contents of this thesis, in full or in parts, have not been
submitted to any other Institute or University for the award of any degree or
diploma.

\vspace*{1.4in}
\hspace*{-0.25in}
\begin{singlespace}
\noindent {\bf Dr.~Shankar~ Balachandran} \\
\noindent Research Guide \\ 
\noindent Assistant Professor \\
\noindent Dept. of Computer Science and Engineering\\
\noindent IIT-Madras, 600 036 \\
\end{singlespace}
\vspace*{0.20in}
\noindent Place: Chennai\\ 
Date:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Acknowledgements
\acknowledgements

I would like to thank everyone who helped me.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abstract

\abstract

\noindent KEYWORDS: \hspace*{0.5em} \parbox[t]{4.4in}{Markov Decision Processes,
Symmetries, Abstraction}

\vspace*{24pt}

Current approaches to solving Markov Decision Processes (MDPs), the de-facto
standard for modeling stochastic sequential decision problems, scale poorly
with the size of the MDP. When used to model real-world problems though,
MDPs exhibit considerable implicit redundancy, especially in the form of
symmetries.  However, existing model minimization approaches do not
efficiently exploit this redundancy due to symmetries. These approaches
involve constructing a reduced model first and then solving them. Hence we
term these as ``explicit minimization'' approaches.

In the first part of this work, we address the question of finding
symmetries of a given MDP. We show that the problem is \emph{Isomorphism
Complete}, that is, the problem is polynomially equivalent to verifying
whether two graphs are isomorphic. Apart from the theoretical importance of
this result it has an important practical application. The reduction
presented can be used together with any off-the-shelf Graph Isomorphism
solver, which performs well in the average case, to find symmetries of an
MDP. In fact, we present results of using NAutY (the best Graph Isomorphism
solver currently available), to find symmetries of MDPs. We next address the
issue of exploiting the symmetries of a given MDP. We propose the use of an
explicit model minimization algorithm called the $G$-reduced image
algorithm that exploits symmetries in a time efficient manner. We present an
analysis of the algorithm and corroborate it with empirical results on a
probabilistic GridWorld domain and a single player GridWorld Soccer domain.
We also present results of integrating the symmetry finding with the
explicit model minimization approach to illustrate an end-to-end approach
for ``Abstraction using Symmetries in MDPs''.

We then note some of the problems associated with this explicit scheme and
as a solution present an approach wherein we integrate the symmetries into
the solution technique implicitly. However, we should select a suitable
solution technique so that the overheads due to integration do not outweigh
the improvements. We validate this approach by modifying the Real Time
Dynamic Programming (RTDP) algorithm and empirically demonstrate
significantly faster learning and reduced overall execution time on several
domains.


\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table of contents etc.

\begin{singlespace}
\tableofcontents
\thispagestyle{empty}

\listoftables
\addcontentsline{toc}{chapter}{LIST OF TABLES}
\listoffigures
\addcontentsline{toc}{chapter}{LIST OF FIGURES}
\end{singlespace}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abbreviations
\abbreviations
 
\noindent 
\begin{tabbing}
xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
\textbf{IITM}   \> Indian Institute of Technology, Madras \\
\textbf{RTFM} \> Read the Fine Manual \\
\end{tabbing}

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Notation

 \chapter*{\centerline{NOTATION}}
 \addcontentsline{toc}{chapter}{NOTATION}
 
 \begin{singlespace}
 \begin{tabbing}
 xxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \kill
 \textbf{$r$}  \> Radius, $m$ \\
 \textbf{$\alpha$}  \> Angle of thesis in degrees \\
 \textbf{$\beta$}   \> Flight path in degrees \\
 \end{tabbing}
 \end{singlespace}
 
 \pagebreak
 \clearpage

%The main text will follow from this point so set the page numbering
%to arabic from here on.
\pagenumbering{arabic}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Background.
 \chapter{Background}
\input{background.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The polyhedral model
 \chapter{The polyhedral model}
\input{polyhedral.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% llvm
 \chapter{Introduction to LLVM}
\input{llvm.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% polly
\chapter{Polly - Pollyhedral optmizations in LLVM}
\input{polly.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% openmp
\chapter{OpenMP Code generation in Polly}
\input{openmp.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% testing
\chapter{Testng with pollybench}
\input{testing.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% gccvsllvm
%\chapter{Comparison of GCC and LLVM}
%\input{gcc_and_llvm.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% Appendices.

\appendix
 
\chapter{Various Tools used in Polyhedral community}
\section{Isl}
\section{Polylib}
\section{Cloog}
\section{PluTo}
\section{Piplib}

\chapter{Doing Projects - The Open Source Way}
 
 Just put in text as you would into any chapter with sections and
 whatnot.  Thats the end of it.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% List of papers

\chapter*{Publications}
\vspace{-0.3cm}

\begin{enumerate}
\item S. M. Narayanamurthy and B. Ravindran (2007). \newblock
  Efficiently Exploiting Symmetries in Real Time Dynamic Programming. \newblock {\em
  IJCAI 2007, Proceedings of the 20th International Joint Conference on
  Artificial Intelligence}, pages 2556--2561.
\end{enumerate}

%\nocite{bellman, Amarel:1968, manning, knoblock90learning,
%crawford92theoretical, Barto:rtdp, Ravindran:proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography.
\pagebreak
\begin{singlespace}
  \begin{small}
	\bibliography{refs}
  \end{small}
\end{singlespace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
