My stay at IIT Madras was enriched by the timely guidance and help from a large number of people. I wish to express my immense grattitude to the
following people who haven helped me in various ways during my stay here:

\emph
{
My giude , Prof. Kamala Krithivasan, for her invaluable giudance and support without which it would not have been possible to put together this report.
She has been a continuous source of motivation and encouragement. I am especially gratful for immence amount of patience she has had through all 
the stages of my work. Also I would like to thank her for keeping  the Theoretical Computer Science (TCS-II) Lab a ver lively and homely place. 
The wonderful atmosphere encourages me to come and spend a lot of time in the lab. The lab is an informal place to work in. I have really enjoyed my time 
here.\\\\
Our Head of the Department, Prof. Siva Ram Murthy C, for all encouragements and support he has given me to do my project.\\\\
Our Faculty advisor Dr. Madhu Mutyam, Project co-ordinator Dr. Narayanaswamy N, for their timely advices and support to complete my project in time.\\\\
Dr. Jaylal Sharma, for his valuable suggestions and advices. His renowned course on complexity theory has helped me a lot in my work.\\\\
All faculty members of our departmet, for their help and encouragements.\\\\
Ms.Saradhambal and Ms.Radhaikumar, for patiently listening to my troubles every time.\\\\
Department librarians Ms.Prema and Mr.Balasundaram, for all their help.\\\\
Ms.Lisha Varghese, my under graduate teacher, for her constant support, encouragements and valuable suggestions without which it would not have been 
do my work with this much concentration and interest. \\\\
My labmate Ajeesh Ramanujan, for all the help and support. He helped me a lot in my project work and in the report writing.\\\\
My dear friends Jethin, Pranav, Shyam, Sirosh, Aravind, Jobin and Abhijith, for making my stay at IIT memorable.\\\\
My B.Tech classmates Arun Babu and Subin George, for taking a special care about me during the entire period of my stay in Chennai.\\\\
Amma, Valyammachi, Ammavan, Ammayi, Kannan, Chechi, Chettan and Sreekuttan, for always ensuring that I didnt feel lost here, for their tender concern
 they have for me, and most of all, for their unwavering belief in me, at all times.\\\\
My best friends Raghesh, Balagopal, Jyothikrishna and Sunitha, for making sure I never really felt lonely while I was here. I thank them most of all, for always boosting my spirits, especially in the most trying of times. I could not have asked for a better group of friends.
}









