 \label{chap:intro}
 
\section{Theoretical Computer Science}
Language is a systematic means of communicating by the use of sounds or conventional symbols. All language have its own alphabet from which 
words are formed. A meaningful collection of words is known as a sentence. For every lanuage there is some rules for sentence formation. ie.,
Grammar is the system of relationships between elements of the sentence that links the ‘sounds’ to the ‘meanings’.

A formal language is a set of words, i.e. finite strings of letters, symbols, or tokens. The set from which these letters are taken is called the 
alphabet over which the language is defined. A formal language is often defined by means of a formal grammar.
Formal languages are studied in computer science and linguistics; the field of formal language theory studies the purely syntactical aspects 
of such languages.

A formal grammar (sometimes simply called a grammar) is a set of rules for forming strings in a formal language. The rules describe how to form strings from the language's alphabet that are valid according to the language's syntax. A grammar does not describe the meaning of the strings or what can be 
done with them in whatever context—only their form.

\subsection{Importance}
\subsection{Applications}
%Grammaras and Automata
\section{Transducers}
\subsection{Tree Transducers}
\subsubsection{Different Types}
\subsection{Applications}
%Applications
\section{Insertion Deletion Systems}
\subsection{In Strings}
%Power
\subsection{In Trees}
\subsubsection{Difference when compared to string insertion}
\section{Motivation for this Work}
\section{Contributions of the Report}
\section{Outline of the Report}

