\documentclass[a4paper,12pt]{article}
%\usepackage{fancyvrb,relsize}
\usepackage[small,compact]{titlesec}
\usepackage[pdftex]{graphicx}
\usepackage{listings}
\lstset{language=C}
%\usepackage[margin=3.50cm]{geometry}
\linespread{1.5}
\setlength{\parindent}{10pt}
\setlength{\parskip}{0.85ex plus 0.65ex minus 0.3ex}
\sloppy
\setlength{\oddsidemargin}{0in} \setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in} \setlength{\textheight}{9.5in}
\setlength{\topmargin}{-0.65in}

\begin{document}

\begin{center}
{\bf {\LARGE Chapter 1}\linebreak\linebreak{Background}}
\linebreak
\linebreak
\end{center}

\section{Auto Parallelization}
\section{The Polyhedral Model}

Polyhedral optimizations are used for many kind of memory access optimization by
looking into the memory access pattern of any piece of code. Any kind of classical
 loop optimization techniques like tiling can be used for this purpose. 


\section{LLVM}
\section{Polly}
The framework discussed in this document is implemented using Polly[polly],
an open source[licence] compiler  optimization framework that uses a mathematical
 representation, the polyhedral model, to represent and transform loops and other
 control flow structures. It is an effort towards achieving autoparallelism in programs.
 The transformations are being implemented in LLVM(Low level virtual machine). 
Polly can detect parallel loops, issue vector instructions and generate OpenMP code(focus of 
this document) corresponding to those loops. Polly try to expose more parallelism
with the help of polyhedral model. A loop which does not look parallel can be transformed
to a parallel loop and these can be vectorized or parallelize using OpenMP.

(grosser video 1.09m)

\section{Manual OpenMP Code Generation}

\end{document}
